---
title: "HW5 ECG"
author: "Burak Şimşek - IE582 - Fall 2020"
date: "2/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###HOMEWORK 5 ECG

```{r libraries, warning=FALSE ,message=FALSE}
#loading libraries
library(caret)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(data.table)
library(glmnet)
library(ROSE)
```

#PART A

After receiving data, logistic regression model fit found and then prediction of test class values found by using test raw data.

```{r read data, warning=FALSE ,message=FALSE}
#reading dataframes
ecg_train <- read.table("https://raw.githubusercontent.com/BU-IE-582/fall20-buraksimsekkk/master/HW5/ecgTRAIN")
ecg_test <- read.table("https://raw.githubusercontent.com/BU-IE-582/fall20-buraksimsekkk/master/HW5/ecgTEST")


```

```{r data manuplation, warning=FALSE ,message=FALSE}
#reading dataframes

x_ecg_train <- ecg_train[,2:ncol(ecg_train)] %>% as.matrix()
y_ecg_train <- ecg_train[,1]
x_ecg_test  <- ecg_test[,2:ncol(ecg_test)] %>% as.matrix()
y_ecg_test  <- ecg_test[,1]


set.seed(50)
model <- cv.glmnet(x_ecg_train, y_ecg_train, family='binomial',nfolds=10)


ecg_predicted_test   <- predict(model, x_ecg_test,s='lambda.min',type = "class")


cm_test <- confusionMatrix(as.factor(ecg_predicted_test), as.factor(y_ecg_test))


balanced_accuracy_test  <- cm_test[["byClass"]][["Balanced Accuracy"]]

#printing confusion matrix results of model1
print(cm_test$table)
```


```{r confusion matrix1, warning=FALSE ,message=FALSE}
#printing confusion matrix results of model1
print(balanced_accuracy_test)

```
The model1 gives 79,07% accuracy.

#PART B


```{r model lambda.min, warning=FALSE ,message=FALSE}
model[["lambda.min"]]

```


```{r model lambda.graph, warning=FALSE ,message=FALSE}

plot(model[["lambda"]])

```

#PART C

```{r time diff, warning=FALSE ,message=FALSE}
#reading dataframes

time_diff_train <- matrix(nrow = nrow(x_ecg_train), ncol = ncol(x_ecg_train)-1)
time_diff_test <- matrix(nrow = nrow(x_ecg_train), ncol = ncol(x_ecg_train)-1)

for (i in 2:96) {
  
time_diff_train[,i-1] <- x_ecg_train[,i] - x_ecg_train[,i-1]
time_diff_test[,i-1]  <- x_ecg_test[,i] - x_ecg_test[,i-1]

}

#time_dif <- time_dif[-90,]
#x_ecg_test <- x_ecg_test[-90,]
#x_ecg_train <- x_ecg_train[-90,]
#y_ecg_test <- as.matrix(y_ecg_test)
#y_ecg_test <- y_ecg_test[-90,]
#y_ecg_train <- as.matrix(y_ecg_train)
#y_ecg_train <- y_ecg_train[-90,]


set.seed(50)
model2 <- cv.glmnet(time_diff_train, y_ecg_train, family='binomial',nfolds=10)


diff_ecg_predicted_test   <- predict(model2, time_diff_test, s='lambda.min' , type = "class")


cm_test2  <- confusionMatrix(as.factor(diff_ecg_predicted_test), as.factor(y_ecg_test))


balanced_accuracy_test2  <- cm_test2[["byClass"]][["Balanced Accuracy"]]

#printing confusion matrix results of model2
print(cm_test2$table)
```
The model gives 79,07% accuracy.


```{r confusion matrix2, warning=FALSE ,message=FALSE}
#printing confusion matrix results of model2
print(balanced_accuracy_test2)

```

The model2 gives 79,68% accuracy.

#PART D

```{r model2 lambda.min, warning=FALSE ,message=FALSE}
model2[["lambda.min"]]

```

```{r model2 lambda.graph, warning=FALSE ,message=FALSE}

plot(model2[["lambda"]])

```

The model2 result is similar to model1 since there is similar relation, which can be interpred by logistic regression ,between data values and their intervals. Thus, We  obtain very close test accuracy.

#PART E

```{r combined data, warning=FALSE ,message=FALSE}
#reading dataframes

comb_train <- cbind(x_ecg_train, time_diff_train)
comb_test  <- cbind(x_ecg_test, time_diff_test)

set.seed(50)
model3 <- cv.glmnet(comb_train, y_ecg_train, family='binomial',nfolds=10)


comb_ecg_predicted_test   <- predict(model3, comb_test, s='lambda.min' , type = "class")


cm_test3  <- confusionMatrix(as.factor(comb_ecg_predicted_test), as.factor(y_ecg_test))


balanced_accuracy_test3  <- cm_test3[["byClass"]][["Balanced Accuracy"]]


#printing confusion matrix results of model3
print(cm_test3$table)
```

```{r confusion matrix3, warning=FALSE ,message=FALSE}
#printing confusion matrix results of model3
print(balanced_accuracy_test3)

```

The model3 gives 73,52% accuracy.

#PART F

```{r model3 lambda.min, warning=FALSE ,message=FALSE}
model3[["lambda.min"]]

```

```{r model3 lambda.graph, warning=FALSE ,message=FALSE}

plot(model3[["lambda"]])

```
The accuracy of model 3 is lower compare to other two model. This might be caused due to overfitting since data values and their differances (intervals) can give similar information.It is better to use either model1 or model2.