---
title: "HW5"
author: "Burak Şimşek - IE582 - Fall 2020"
date: "2/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###HOMEWORK 5 uWaveGesture

```{r libraries, warning=FALSE ,message=FALSE}
#loading libraries
library(ggplot2)
library(dplyr)
library(tidyverse)
library(data.table)
library(caret)
library(skimr)
```

```{r read data, warning=FALSE ,message=FALSE}
#reading dataframes
x_train <- read.table("https://raw.githubusercontent.com/BU-IE-582/fall20-buraksimsekkk/master/HW5/uWaveGestureLibrary_X_TRAIN")
y_train <- read.table("https://raw.githubusercontent.com/BU-IE-582/fall20-buraksimsekkk/master/HW5/uWaveGestureLibrary_Y_TRAIN")
z_train <- read.table("https://raw.githubusercontent.com/BU-IE-582/fall20-buraksimsekkk/master/HW5/uWaveGestureLibrary_Z_TRAIN")
x_test  <- read.table("https://raw.githubusercontent.com/BU-IE-582/fall20-buraksimsekkk/master/HW5/uWaveGestureLibrary_X_TEST")
y_test  <- read.table("https://raw.githubusercontent.com/BU-IE-582/fall20-buraksimsekkk/master/HW5/uWaveGestureLibrary_Y_TEST")
z_test  <- read.table("https://raw.githubusercontent.com/BU-IE-582/fall20-buraksimsekkk/master/HW5/uWaveGestureLibrary_Z_TEST")

```


#PART A

Since there are x, y,z dimentions on both train and data, we need to combine them.

```{r data manupulation , warning=FALSE ,message=FALSE}
#column wide binding of x, y,z data

train <- cbind(x_train, y_train, z_train)
test  <- cbind(x_test, y_test, z_test)


#changeing train and test data colomn name to avoid duplications.
train_features <- colnames(train) %>% make.names(unique = TRUE)
test_features  <- colnames(test) %>% make.names(unique = TRUE)

colnames(train) <-train_features
colnames(test)  <-test_features

```

To fit a model we need to scale the data. 

```{r data manupulation2 , warning=FALSE ,message=FALSE}
#Scaling both train and test data.

train[,-1] <- scale(train[,-1])
test[,-1] <- scale(test[,-1])

#removing target value from train and test data
train_data <- train[,-1]
test_data  <- test[,-1]

#selecting target values of test data
class_test <- test$V1 %>% as.factor

```

After all data manupulations completed, fit model can be created.

```{r knn , warning=FALSE ,message=FALSE}
#making knn model with 3 repeatation with 10-fold cross validation for 5 different k valee (5, 7, 9, 11, 13)

set.seed(50)
n_folds <- 10
ctrl <- trainControl(method="repeatedcv",repeats = 3, number = n_folds) #,classProbs=TRUE,summaryFunction = twoClassSummary)
knnFit <- train(V1 ~ ., data = train, method = "knn", trControl = ctrl, preProcess = c("center","scale"), tuneLength = 5)
print(knnFit)
```

Five different k measure is tested. k=5 is minimize the error of a 10-fold cross validation with 0,60 RMSE valeu.
Thus, we will use k=5 for part b and c.

#PART B

```{r knn prediction, warning=FALSE ,message=FALSE}
#Making prediction by taking k=5

pred <- predict(knnFit, test_data) 
pred <-  round(pred,0) %>% as.factor

cm <- confusionMatrix(pred, class_test)

print(cm$table)
print(cm[["overall"]][["Accuracy"]])
print(cm[["byClass"]])

```

The confusion matrix of model is shown in the above table. The general accuracy of model gives 88.27%
The separate class accracy can also be seen.



#PART C
If we train model according to final_dist = w1distx, + w2disty + w3*distz, it will change the accuray of model. This method might be reasonable if one of the distance effect is higher than other. We can handle this unbalancing by weighting distancs.





